---
title: "ANALYSIS REPORT"
author: "XXX"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    includes:
    code_folding: "hide"
    toc: true
    toc_float: false
    number_sections: false
bibliography:
- r-references.bib
  # repro:
#   data:
#     data: data.csv
#     
#   #scripts: clean.R
---

### Setup {-}
  
```{r setup, results='hide', message=FALSE, warning=FALSE}

#this is to check to see if package are installed and if not to install them or just load them if they are installed!
if(!require(pacman)) {install.packages(c("pacman", "remotes"))}

if(!require(papaja)) {remotes::install_github("crsh/papaja")}

pacman::p_load(here, tinylabels, apaTables, tidyverse, gtsummary, car, GGally, MASS, rcompanion, moments, utils, sjPlot, interactions, kableExtra, report, papaja, bookdown)

p_load_gh("crsh/papaja@devel")


#get relative path
path =  here("RESSOURCES") #this is really important ! you juste need to put the name of YOUR folder and here() will find the relative path for you ! 
#or path =  dirname(rstudioapi::getActiveDocumentContext()$path) in R
setwd(path) #set working directory

# load data
base  <- read.csv("data.csv")


session = utils::sessionInfo() # get session informations

#opts_chunk$set(echo = F, message=F, warning=F) # set echo F for all

```

<!-- This file was automatically created via `the Repro package (version 0.1.0)` using  `r session$R.version$version.string` -->



```{r options, results='hide', message=FALSE, warning=FALSE}
options(scipen = 666, warn=-1, contrasts=c("contr.sum","contr.poly"), knitr.kable.NA = '')  #remove scientific notation # remove warnings #set contrasts to sum ! 
set.seed(666) #set random seed

`%notin%` <- Negate(`%in%`)

# panderOptions('knitr.auto.asis', FALSE) #remove auto styling
```


$~$
$~$

### Introduction
Blabla citation in parentesis [@james_1890] blbabla, or citation in text @james_1890

$~$
$~$

#### Clean up 

$~$
$~$
```{r clean}
base$Priming = as.factor(base$Priming); base$Product = as.factor(base$Product); 
```

$~$
$~$

#### Demographics

$~$
$~$
```{r demographics}
base[c("Age", "Gender", "Profession")] %>%  tbl_summary(statistic = list(all_continuous() ~ "{mean} ({sd})", all_categorical() ~ "{n} / {N} ({p}%)"),)  %>% modify_caption("**Table 1. Demographics **") %>%bold_labels()
```
$~$
$~$

### Inspections

$~$
$~$
Inspect all distributions
```{r plots, fig.align="center", message=FALSE, warning=FALSE}
base %>% 
  keep(is.numeric) %>%
  gather() %>%
  ggplot(aes(value)) +
  facet_wrap(~key, scales = "free") + 
  geom_histogram()

#or GGally
```
$~$
$~$

Inspect distributions by groups with violin plots!
```{r violin, fig.align="center", message=FALSE, warning=FALSE}
ggplot(base, aes(x=Priming, y=Age)) +
  geom_violin(trim=FALSE, fill='#A4A4A4', color="darkred")+
  geom_boxplot(width=0.1) + theme_minimal()
```

$~$
$~$

### Statistics

$~$
$~$

#### Reccomendation

```{r Report_Regression}
base_recco = filter(base, id %notin% c("13594")) #remove huge outliers, see appendix

final_recco = lm(Reccomend ~Priming*Product + Behav_enviro + Product:Behav_enviro,  data = base_recco)

apa_lm <- papaja::apa_print(anova(final_recco))
papaja::apa_table(apa_lm$table,  caption = "Anova table for Reccomendation.", label =NULL)


sjPlot::plot_model(final_recco)

```
$~$
$~$
```{r GLM int, fig.align = 'center'}
interactions::interact_plot(final_recco, pred = Behav_enviro, modx = Product, interval = TRUE)
```
\

$~$
$~$

### Conclusions

$~$
$~$

Example for reporting :\

Product (`r apa_lm$full_result$Product`) and Behav_enviro (`r apa_lm$full_result$Behav_enviro`) affected reccomendation. However, the effect of Behav_enviro differed by Product, `r apa_lm$full_result$Product_Behav_enviro`.

$~$
$~$


### Appendix

$~$
$~$

#### Visual check normality

```{r annexes, fig.show="hold", out.width="50%", fig.align = 'center'}

plotNormalHistogram(base$Reccomend, main = "Reccomendation", sub = paste("skewness =", round(skewness(base$Reccomend, na.rm = TRUE),2)))

```

$~$
$~$


#### Diagnostics Plots for GLM Reccomendation with outliers
```{r diag GLM Recco, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold", out.width="50%"}

modrecco = lm(Reccomend ~ Priming*Product*Decision_mode +Age+ Priming*Product*Behav_enviro, data = base)

plot(modrecco, c(1:2,4), labels.id = base$id )
```

$~$
$~$

#### Diagnostics Plots for GLM Reccomendation with outliers
```{r diag GLM Recco2, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold", out.width="50%"}

modrecco = lm(Reccomend ~Priming*Product*Decision_mode +Age+ Priming*Product*Behav_enviro, data = base_recco)

plot(modrecco, c(1:2,4), labels.id = base_recco$id )
```

$~$
$~$

#### Model selection for Reccomendation
```{r diag GLM MSrecco, echo=FALSE, message=FALSE, warning=FALSE}

MS = MASS::stepAIC(modrecco, direction = "both", trace = FALSE)

x = kable(attributes(MS$anova)$heading, format="html") %>%
    kable_styling(latex_options = "HOLD_position", position = "center", full_width = F) %>%  row_spec(1, bold=T,align='c') %>%  row_spec(5, bold=T)
gsub("<thead>.*</thead>", "", x)


kable(MS$anova, format="html") %>%
    kable_styling(latex_options = "HOLD_position", position = "center", full_width = F) %>%  row_spec(0, bold=T,align='c')

# sjPlot::plot_model(final_recco, type = "pred")
```
$~$
$~$

# Packages

$~$
$~$
```{r, comment=NA}
report::report_packages()
```

$~$
$~$

# References
$~$
$~$
```{r, include = F}
papaja::r_refs(file = "r-references.bib")
my_citation <- cite_r(file = "r-references.bib", )
# `r_refs()` creates a BibTeX file containing citations for R and all currently loaded packages.
# `cite_r()` takes these citations and turns them into readily reportable text.
# `my_citation` now contains the following text that you can use in your document: `r my_citation`

```
$~$
$~$
