---
title: "W12 TP"
author: "Lucile Favero"
date: "12/6/2021"
output: 
  pdf_document:
    toc: true
    toc_depth: 1
    number_sections: true
    df_print: kable
---
# Aim
**Report how bone marrow transplant survival times relates to graft versus host disease (GHVD)**

# Preprocessing 
## Load library and data
```{r}
library(GGally)       # for ggpairs
library(ggfortify)    # for autoplot
library(ggplot2)      # for ggplot
library('MASS')       # for the glm model selection 
library(glmnet)
library(patchwork)
library(ISwR)
library(survival)
library(survminer)

d<- graft.vs.host
```
## Understand datas
 
 >"The gvhd data frame has 37 rows and 7 columns. It contains data from patients receiving a nondepleted >allogenic bone marrow transplant with the purpose of finding variables associated with the development of >acute graft-versus-host disease."
 
```{r}
str(d)
summary(d)
```
Transform into factor the variables : type, preg, gvhd, dead
```{r}
d$type <-as.factor(d$type)#type of leukaemia coded 1: AML, 2: ALL, 3: CML for acute myeloid, acute lymphatic, and chronic myeloid leukaemia.
d$preg<-as.factor(d$preg)# indicating whether donor has been pregnant. 0: no, 1: yes.``
levels(d$preg)<-c("no","yes")
d$gvhd<-as.factor(d$gvhd)# graft-versus-host disease, 0: no, 1: yes
levels(d$gvhd)<-c("no","yes")
d$dead <- as.factor(d$dead) # a numeric vector code, 0: no (censored), 1: yes
levels(d$dead)<-c("no","yes")
str(d)
summary(d)
```
Plot the ggpairs:
```{r}
ggpairs(d, aes(color=gvhd, alpha = 0.3))
```

# Modelization 

```{r}
surv<-Surv(d$time,d$dead==1)
```

## A simple model
We set a model depending only on gvhd
```{r, warning=FALSE}
m1<-coxph(data=d,surv~ gvhd)
summary(m1)
```

The variable is significant for this model.
```{r}
m1.diag<-cox.zph(m1)
m1.diag
p1<-ggsurv(survfit(m1))+ylim(0,1)
p2<-ggcoxdiagnostics(m1,hline=FALSE)+ geom_smooth()
ggcoxzph(m1.diag)
(p1 | p2)
```
The p-value is larger than 5%, so the null hypothesis of proportional hazards is rejected.
However, the second plot has a correct regression line and the in the third, all the points are in the CI.


## Add more variables
```{r, warning=FALSE}
m2<-coxph(data=d,surv~ gvhd+rcpage+ donage+ type+ preg +index )
summary(m2)
```

The two variables significant for this model are gvhd and type.
```{r, warning=FALSE}
m2.diag<-cox.zph(m2)
m2.diag
p1<-ggsurv(survfit(m2))+ylim(0,1)
p2<-ggcoxdiagnostics(m2,hline=FALSE)+ geom_smooth()
p3<-ggcoxzph(m2.diag)
p1 | p2
p3
```
From the second plot, we see that the linear prediction is not respected.
The p-value is larger than 5%, so the null hypothesis of proportional hazards is rejected.
In the last plot all the points are in the CI.
## only gvhd and type
```{r, warning=FALSE}
m3<-coxph(data=d,surv~ gvhd+type )
summary(m3)
```
The two variables are significant. 
```{r, warning=FALSE}
m3.diag<-cox.zph(m3)
m3.diag
p1<-ggsurv(survfit(m3))+ylim(0,1)
p2<-ggcoxdiagnostics(m3,hline=FALSE)+ geom_smooth()
p3<-ggcoxzph(m3.diag)
p1 | p2
p3
```
From the second plot, we see that the linear prediction is not respected.
The p-value is lower than 5%, so the null hypothesis of proportional hazards is not rejected.
In the last plot all the points are in the CI.
# Version of R used 
```{r, echo = FALSE}
sessionInfo()
```
