---
title: "DAP Project"
author: "Daniel Granja"
date: "10/12/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#this is to check to see if package are installed and if not to install them or just load them if they are installed!
if(!require(pacman)) {install.packages("pacman")}

pacman::p_load(here, ggplot2, rstatix, ordinal,stargazer, RVAideMemoire,sjPlot, HH, likert, dplyr,sure)


#get relative path
path =  here("DANIEL") #You put the name of YOUR folder and here() will find the relative path for you ! 
setwd(path) #set working directory

 
#data scanning 
Raw_data <- read.csv2("Raw_data.csv")
```


Introduction
============

Context of the project
----------------------

According to the American Psychological Association (APA), emotion is defined as “a complex reaction pattern, involving experiential, behavioral and physiological elements.”
Researchers have been studying the impact of emotions on cognitive functions, such as memory. Studies have shown that emotional activation, and specifically emotions of negative valence, favours the retrieval processes of associative memory for single items, but impairs it for associated items or contexts (Schmidt, Patnaik & Kensinger, 2010).
It's believed that the neurobiological systems sustaining high emotional arousal and memory are linked, especially through the adrenal stress hormones, itself mediated by the amygdala activity (McGaugh, 2013).
In the study the dataset came from, Riegel et al. (2020) used words to elicit emotions (disgust, fear or neutral). Each pair displayed the same emotion and when showed together, they had to imagine an interaction between them, forcing the creation of a meaningful association. To be sure that the displayed words showed the correct emotion, affective ratings were asked as controls. Each participant had to rate each pair on a Likert scale (1 to 7) of disgust and fear.
In this project, we are going to use this dataset to analyse and compare three statistical methods. We will compare a repeated measures ANOVA, a ordinal polytomous logistic regression and a linear mixed model. 


Data description
----------------
Here is a description of the different variables:

* **Subjects**: the subject's number in the experiment.

* **Wordpairs**: the wordpairs presented.

* **Emotion**: the emotion elicited by the wordpairs (disgust, fear or neutral).

* **Gender**: the gender of the subjects (man or woman).

* **Disgust**: a Likert scale (1 to 7) of the subjects’ emotional subjective feeling of disgust.

* **Fear**: a Likert scale (1 to 7) of the subjects’ emotional subjective feeling of fear.


```{r, echo=FALSE}
#Creating a working dataset
#Giving nice names to variables

subjects<-Raw_data$ï..code
wordpairs<-Raw_data$wordpair
emotion<-Raw_data$emotion
gender<-Raw_data$sex
disgust<-Raw_data$disgust
fear<-Raw_data$fear

#Creating a dataset
Data<-data.frame(subjects,wordpairs,emotion,gender,disgust,fear)

#Adapting the variables
Data$emotion<-as.factor(Data$emotion)
levels(Data$emotion) <- c("Disgust", "Fear","Neutral")
Data$gender<-as.factor(Data$gender)
levels(Data$gender) <- c("Women", "Men")
Data$disgust<-as.factor(Data$disgust)
Data$fear<-as.factor(Data$fear)
```

Data exploration (Dataset)
----------------

```{r basic explor, comment = NA}
str(Data)


summary(Data)
```
Data exploration (Graphs) 
-------------------------

```{r}

## Prepare data <- **need to specify emotion$disgust + the other ratings**
a=9349
Item <- "Disgust"
one <- (6611/a)*100
two <- (972/a)*100
three <- (706/a)*100
four <- (432/a)*100
five <- (288/a)*100
six <- (174/a)*100
seven <- (166/a)*100
df <- data.frame(Item, one, two, three, four, five, six, seven)
## Rename Cols
df <- df %>%
rename("1 (low)" = one, "2" = two, "3" = three, "4" = four, "5" = five, "6"= six, "7 (high)"= seven)

## Pretty Plot
plot(likert(summary = df))+ guides (fill = guide_legend(title="Subjective evaluation")) + labs(title = "Subjective evaluation of disgust across all pairs of words")+ labs(y = "Percentage of Likert answers") + theme(plot.title = element_text(hjust = 0.5))+ theme(plot.title = element_text(size = rel(1.5))) + theme( legend.box.background = element_rect(), legend.box.margin = margin(6, 6, 6, 6),legend.position = "top")

##Disgust depending on emotion presented
ggplot(Data, aes(emotion,disgust,colour=gender))+geom_jitter()+
  labs(title = "Distribution of disgust evaluation", x = "Pairs of words' emotions", y = "Likert scale of subjective disgust feeling") + theme(plot.title = element_text(hjust = 0.75)) + labs(colour = "Gender")+ theme(plot.title = element_text(size = rel(2)))
```

```{r modeling}
#logistic ordinal regression
m1.1 = MASS::polr(disgust ~ emotion + gender, data = Data)
m1.2 = MASS::polr(disgust ~ emotion * gender, data = Data)

m2.1 = clmm(disgust ~ emotion + gender + (1|subjects), data = Data)
m2.2 = clmm(disgust ~ emotion * gender + (1|subjects), data = Data)

m3.0 = clmm(disgust ~ emotion + (emotion | subjects), data = Data)
m3.1 = clmm(disgust ~ emotion + gender + (emotion | subjects), data = Data)
m3.2 = clmm(disgust ~ emotion * gender + (emotion | subjects), data = Data)

anova(m1.1,m1.2)#better
anova(m2.1,m2.2)#better
anova(m3.0,m3.1)#not better


extractAIC(m1.2)
extractAIC(m2.2)
extractAIC(m3.0)
```

```{r best model analysis}
summary(m3.0)

RVAideMemoire::Anova.clmm(m3.0, type = "II") #or type = "III" depending on the context
```
Diagnostic of the moddel
=======================
```{r}
#autoplot.resid(m3.0) ##do not work
#read this : David W.. Hosmer, Lemeshow, S., & Rodney X.. Sturdivant. (2000). Applied logistic regression. New York: Wiley. --> §5
```