---
title: "DAP Project"
author: "Daniel Granja"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    includes:
    code_folding: "hide"
    toc: true
    toc_float: true
    number_sections: false
    df_print: paged
---


```{r setup, warning=FALSE, message=FALSE}
#this is to check to see if package are installed and if not to install them or just load them if they are installed!
if(!require(pacman)) {install.packages("pacman")}

pacman::p_load(here, ggplot2, ordinal, ggfortify, GGally, grid,gridExtra, plyr, sjPlot, likert, vtable, brant, Hmisc, MASS, kableExtra, gtsummary, PResiduals)

#get relative path
path =here("DANIEL")

#load data 

Raw_data <- read.csv2("Raw_data.csv")
```


## Introduction


### Context of the project


According to the American Psychological Association (APA), emotion is defined as “a complex reaction pattern, involving experiential, behavioral and physiological elements.”
<br>

Researchers have been studying the impact of emotions on cognitive functions, such as memory, and have shown that emotional activation, and specifically emotions of negative valence, favours the retrieval processes of associative memory for single items, but impairs it for associated items or contexts (Schmidt, Patnaik & Kensinger, 2010).
<br>

It's believed that the neurobiological systems sustaining high emotional arousal and memory are linked, especially through the adrenal stress hormones, itself mediated by the amygdala activity (McGaugh, 2013).
<br>

In the study the dataset came from, Riegel et al. (2020) used words to elicit emotions (disgust, fear or neutral). Each pair displayed the same emotion and when showed together they had to imagine an interaction between them, forcing the creation of a meaningful association. To be sure that the displayed words showed the correct emotion, affective ratings were asked as controls. Each participant had to rate each pair on a Likert scale (1 to 7) of disgust and fear.
<br>

In this project, we are going to use part of this dataset to analyse and compare three statistical methods. We will compare (a) a simple logistic ordinal regression, (b) a simple mixed-effect model only considering the subjects' repetition as a random effect, and (c) a more complete mixed-effect model with random slopes allowing fixed effects to vary for each emotion. We follow the 'Keep it maximal' procedure from Barr et al. (2013) to find a model that best fits our data, considering how the emotion elicited in the word pairs and the gender impact the subjective ratings of disgust by the participants.


### Data description
Here is a description of the different variables:

* **Subjects**: the subject's number in the experiment.

* **Wordpairs**: the word pairs presented.

* **Emotion**: the emotion elicited by the word pairs (disgust, fear or neutral).

* **Gender**: the gender of the subjects (man or woman).

* **Disgust**: a Likert scale (1 to 7) of the subjects’ emotional subjective feeling of disgust while judging the word pairs. 


```{r}
#Creating a working dataset
#Giving nice names to variables

subjects<-Raw_data$code
wordpairs<-Raw_data$wordpair
emotion<-Raw_data$emotion
gender<-Raw_data$sex
disgust<-Raw_data$disgust
fear<-Raw_data$fear

#Creating a dataset
Data<-data.frame(subjects, wordpairs, emotion, gender, disgust)
```

```{r}
#Adapting the variables
Data$emotion<-as.factor(Data$emotion)
levels(Data$emotion) <- c("Disgust", "Fear","Neutral")
Data$gender<-as.factor(Data$gender)
levels(Data$gender) <- c("Women", "Men")
Data$disgust<-as.factor(Data$disgust)
```

## Data overview {.tabset}

### Structure
```{r basic explor, comment = NA}
str(Data)
```

### Summary
```{r comment = NA}
# st(Data) # replacded by that so its nicer
Data[c("emotion", "gender", "disgust")] %>%  tbl_summary()  %>% modify_caption("**Summary statistics**") %>% bold_labels()
```
As one can see in this table, the distribution of emotion through the word pairs and the gender is equivalent. 
However, the distribution of the disgust response has a floor effect, more than 80% of it being 1 and 2.



## Visual exploration 

```{r Likert representation}
## Prepare data for plot <- **need to specify emotion$disgust + the other ratings**
Newdata<-data.frame(Data$emotion,Data$disgust)

df = prop.table(table(Newdata$Data.emotion, Newdata$Data.disgust), margin = 1)*100

#this is just ot make it fit into Likert
df = as.data.frame.matrix(df)
df$Item <- row.names(df)
rownames(df) = NULL
df = df[,c(8,1,2,3,4,5,6,7)]
colnames(df) = c("Item", "1 (low)","2","3","4","5","6","7 (high)")
```


```{r}
# Pretty Plot
plot(likert(summary = df))+ 
  guides(fill = guide_legend(name="Subjective evaluation"))+ 
  labs(title = "Subjective evaluation of disgust across all pairs of words")+
  labs(y = "Percentage of Likert answers") +
  labs(x="Word pairs' emotions") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.title = element_text(size = rel(1.5))) +
  theme(legend.box.background = element_rect(), legend.box.margin = margin(6, 6, 6, 6), legend.position = "right")
```

This figure shows the distribution of the subjective evaluation of disgust across all pair of words, depending of the emotion elicited by the words. As one can see, there still is a floor effect, but it seems less strong for the disgust emotion, as one can expect since the emotion elicited corresponds to the emotion rated. However, this floor effect can be explained by the difficulty of identification of disgust (Yoder et al., 2016). Since it's difficult to specifically recognize disgust, mixed to the small arousal of some word pairs, there is a possible existence of a bias in the ratings. Another representation including the gender can be found in Appendix 1. 
<br>

## Modeling

Even if we have seen in our exploration that there seems to be a floor effect, we are going to do a diagnostic of a linear model to certify the unfit of this model to analyse our data.

```{r Basic analysis, include=FALSE}
disgust2<-as.numeric(disgust)
m0 = lm(disgust2~emotion*gender, data=Data)
plot(m0)
```

```{r normal analysis of resids}
sjPlot::plot_model(m0,type = "diag")[[3]]
```

As expected, the residuals distribution does not follow a normal curve. Therefore, we cannot do a classic linear regression to analyse this data. Alternative techniques have to be investigated. (Find more diagnostic graphs in the Appendix 2.)

Three possible methods are being tested: 
(a) a simple logistic ordinal regression, (b) a simple mixed-effect model only considering the subjects' repetition as a random effect, and (c) a more complete mixed-effect model with random slopes allowing fixed effects to vary for each emotion.


```{r, cache=TRUE}
m1.2 = clm(disgust ~ emotion * gender, data = Data) #Model A
m2.2 = clmm(disgust ~ emotion * gender + (1|subjects), data = Data) #Model B
m3.2 = clmm(disgust ~ emotion * gender + (emotion | subjects), data = Data) #Model C
```

After computing the full model for each random structure, we can use the Akaike information criterion to estimate the prediction error of each model, and thereby their relative quality. 
```{r, cache=TRUE}
a1<- extractAIC(m1.2)
a2<- extractAIC(m2.2)
a3<- extractAIC(m3.2)

c1<-cbind(a1,a2,a3)
r1<-c1[1,]
r2<-c1[2,]
c2<-as.data.frame(cbind(r1,r2))
rownames(c2) <- c("No random structure", "Simple random strucure","Full random structure")
colnames(c2) <- c("Equivalent degrees of freedom", "Akaike Information Criterion")
c2
```

As shown in this table, we can conclude that the model that seems to fit the best our data is the full random structure model, a mixed-effect model with random slopes allowing fixed effects to vary for each emotion, as it has the smaller AIC, and more equivalent degrees of freedom.

We can test explicitly the random slope for emotions' fixed effects to confirm the importance to have this complexity in the model. 

```{r random slope, cache=TRUE, message=FALSE, warning=FALSE}
anova(m2.2, m3.2)
```
It is important to include the random slope for emotions' fixed effects (χ²(5)=138.5, p < 0.001).

This being done, we must now look for the best model using this random structure. According to our initial hypothesis, we want to know if there is an impact of the word pairs' emotion and the gender on the subjective evaluation of disgust. 

```{r best model, cache=TRUE, message=FALSE, warning=FALSE}
#best model
#m3.2 = clmm(disgust ~ emotion * gender + (emotion | subjects), data = Data) #done above
m3.1 = clmm(disgust ~ emotion + gender + (emotion | subjects), data = Data)
m3.0 = clmm(disgust ~ emotion + (emotion | subjects), data = Data)


anova(m3.0, m3.1, m3.2, test='Chisq') 
```
As shown in the test summary above, and also following the AIC estimate, the most parsimonious model is the model without gender effect. 


### Statistical Inference {.tabset}

#### Model Summary
```{r best model summary, comment=NA}
summary(m3.0)
```

Now that we have chosen our best model, we are going to take a deeper look at the estimates.

The estimated model can be written as:
$$Logit(P(Y\le 2))= (-0.5346) - (-2.0596)*emotionFear - (-3.9909)*emotionNeutral$$
$$Logit(P(Y\le 3))= 0.3387 - (-2.0596)*emotionFear - (-3.9909)*emotionNeutral$$
$$Logit(P(Y\le 4))= 1.1585 - (-2.0596)*emotionFear - (-3.9909)*emotionNeutral$$
$$Logit(P(Y\le 5))= 1.8831 - (-2.0596)*emotionFear - (-3.9909)*emotionNeutral$$
$$Logit(P(Y\le 6))= 2.6467 - (-2.0596)*emotionFear - (-3.9909)*emotionNeutral$$
$$Logit(P(Y\le 7))= 3.4641 - (-2.0596)*emotionFear - (-3.9909)*emotionNeutral$$


#### Regression summary table
```{r}
sjPlot::tab_model(m3.0, show.stat = TRUE, show.aic = T, show.reflvl = T, show.est = T, p.style = "stars")
```
The effect of the emotion predictors of fear (OR = 0.13, 95% CI 0.10-0.16, p < 0.001) and neutral (OR = 0.02, 95% CI 0.01-0.02, p < 0.001) are different from disgust. The random slope for emotions' fixed effects has been reported above.
<br>

Being the model that best fits the data doesn't mean that it's a good model, that is why we must check for the explained variance. 
We can see that the Conditional R<sup>2</sup> is 0.528, which means that 52.8% of the variance is explained by both the fixed and random factors. 


### Model visualization {.tabset}

#### Plot estimates
```{r}
plot_model(m3.0, show.values = T, value.offset = .4, title = "Disgust")
```
<br>
The model represented above shows how more a disgust evaluation is high, more the odds of the wordpairs' eliciting disgust is high. It also shows that when the emotion elicited was not disgust, the odds of a high evaluation of disgust are low.

#### Plot marginal effects
```{r, message=FALSE, warning=FALSE}
plot_model(m3.0, type = "pred", terms = "emotion")
```
<br>
The model represented above shows how more a disgust evaluation is high, more the odds of the wordpairs' eliciting disgust is high. It also shows that when the emotion elicited was not disgust, the odds of a high evaluation of disgust are low.


### Diagnostic of the model {.tabset}

Ordinal logistic regressions have four assumptions that need to be controled for when aplied: 

* The dependent variable are ordered.
* One or more of the independent variables are either continuous, categorical or ordinal.
* No multi-collinearity.
* Proportional odds.

The two firsts are respected by the structure of our data (see Visual exploration). We must check for the two others.

#### Residual representation

```{r residual plots}
# Residual-vs-covariate plot and Q-Q plot
m1.3<-polr(disgust ~ emotion, data = Data)
pres <- presid(m1.3)

p1 <- ggplot(data.frame(x = Data$disgust, y = pres), aes(x, y)) +
geom_point(color = "#444444", shape = 19, size = 2, alpha = 0.5) +
ylab("Probability-scale residual")
p2 <- ggplot(data.frame(y = pres), aes(sample = y)) +
stat_qq(distribution = qunif, dparams = list(min = -1, max = 1), alpha = 0.5) +
xlab("Sample quantile") +
ylab("Theoretical quantile")
grid.arrange(p1, p2, ncol = 2) # Figure 1
```
SBS residual plots for the ordinal logistic regression model. Left:
Residual-vs-covariate plot. Right: Q-Q plot of the residuals.


#### No multi-collinearity

```{r}
Data[c("emotion", "gender", "disgust")] %>%  tbl_summary(by=emotion)  %>% modify_caption("**Summary statistics by emotion**") %>%
  add_p() %>% bold_labels()
```
We must check for any collinearity between our predictors. Our final model only has the wordpairs' emotion as predictor, so there is no possible collinearity with other predictors. For the other models, we also included the gender. There is no multicollinearity between gender and emotion (χ²(2, N = 9349) = 0.0006, p > 0.05)

#### Proportional odds

To test for this assumption, we want to do a Brant Test. Unfortunately, we cannot do it directly to our mixed-effect model as the test is limited to ordinal logistic regression models. No alternative has been found during further investigations, even if  the existence of a χ2 score test of the proportional odds assumptions has been noted, without further explanation (Schmidt, 2012).
<br>

However, we can test the model without the random effects and with only the emotion predictor as we may expect that these random effects do not influence the relationship between each pair of outcome groups. 

```{r, cache=TRUE, comment=NA}
brant(m1.3)
```
The test tells us that the assumption is respected, which means that the assumptions for ordinal logistic regression are respected. 

## Conclusion

The present data analysis was aimed to compare three statistical methods and to analyse how the emotion elicited in the word pairs and the gender impact the subjective ratings of disgust by the participants. As shown, the best method is the one that take in consideration the repetition of the subjects and the way each one of them express the emotion elicited. 
<br>

For that, the mixed-effect model with random slopes allowing fixed effects to vary for each emotion and considering the subjects' repetition as a random effect, but not including the gender was the best model tested. 
<br>

In other words, gender doesn't seem significative in the disgust evaluation, but it seems important to control for the way each subject evaluate the emotion. 
<br>

Further investigation can be made with the fear evaluation of this dataset to support or reject this proposition. Fear is usually better recognized by the individuals, and the original dataset has a similar likert scale measuring the fear evaluation for each word pairs. If the same conclusion is found, it's a good indicator that this model is the best to fit this kind of data.  
<br>

Emotion isn't felt by everyone the same way, and every participant evaluated multiple pairs of words. Plus, the ordinal scale of the evaluation created an ensemble of constraints that made us use more sophisticated methods than simple linear methods. However, that mean that the interpretation of these results aren't very straight forward. The coefficients are scaled in terms of logs. Another way to interpret ordinal logistic regression models is to use odd ratios, which are more easy.


 
## Bibliography

Barr, D. J., Levy, R., Scheepers, C., & Tily, H. J. (2013). Random effects structure for confirmatory hypothesis testing : Keep it maximal. Journal of memory and language, 68(3), 10.1016/j.jml.2012.11.001. https://doi.org/10.1016/j.jml.2012.11.001

Emotion – APA Dictionary of Psychology. (s. d.). Retrieved January 4, 2022 from https://dictionary.apa.org/emotion

McGaugh, J. L. (2013). Making lasting memories : Remembering the significant. Proceedings of the National Academy of Sciences, 110(2), 10402‑10407. https://doi.org/10.1073/pnas.1301209110

Riegel, M., Wierzba, M., Wypych, M., Ritchey, M., Jednoróg, K., Grabowska, A., Vuilleumier, P., & Marchewka, A. (2020). Distinct medial-temporal lobe mechanisms of encoding and amygdala-mediated memory reinstatement for disgust and fear (p. 825844). https://doi.org/10.1101/825844

Schmidt, J. (2012). Ordinal Response Mixed Models : A Case Study. 21. Retrieved January 4, 2022 from https://math.montana.edu/grad_students/writing-projects/2012/12schmidt.pdf 

Schmidt, K., Patnaik, P., & Kensinger, E. A. (2011). Emotion’s influence on memory for spatial and temporal context. Cognition and Emotion, 25(2), 229‑243. https://doi.org/10.1080/02699931.2010.483123

Yoder, A. M., Widen, S. C., & Russell, J. A. (2016). The word disgust may refer to more than one emotion. Emotion, 16(3), 301‑308. https://doi.org/10.1037/emo0000118


## Appendix

### 1. Alternative graph
```{r alternate graph}
##Disgust depending on emotion presented
ggplot(Data, aes(emotion,disgust,colour=gender))+geom_jitter(alpha=0.3)+
  labs(title = "Distribution of disgust evaluation", x = "Wordpairs' emotions", y = "Likert scale of subjective disgust evaluation (1-7)") + theme(plot.title = element_text(hjust = 0.75)) + labs(colour = "Gender")+ theme(plot.title = element_text(size = rel(2)))
```

### 2. Linear model diagnostics
```{r}
autoplot(m0)
```


### 3. Package Info
```{r session info, echo = FALSE, comment = NA}
report::report_packages() # more suited here
```
