---
title: "DAP Project"
author: "Daniel Granja"
date: "10/12/2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#this is to check to see if package are installed and if not to install them or just load them if they are installed!
if(!require(pacman)) {install.packages("pacman")}

pacman::p_load(here, ggplot2, rstatix, ordinal,stargazer, RVAideMemoire,sjPlot, HH, likert, dplyr,sure,vtable)

#get relative path
path =  here("DANIEL") #You put the name of YOUR folder and here() will find the relative path for you ! 
#setwd(path) #set working directory

 
#data scanning 
Raw_data <- read.csv2("Raw_data.csv")
```


Introduction
============

Context of the project
----------------------

According to the American Psychological Association (APA), emotion is defined as “a complex reaction pattern, involving experiential, behavioral and physiological elements.”
Researchers have been studying the impact of emotions on cognitive functions, such as memory. Studies have shown that emotional activation, and specifically emotions of negative valence, favours the retrieval processes of associative memory for single items, but impairs it for associated items or contexts (Schmidt, Patnaik & Kensinger, 2010).
It's believed that the neurobiological systems sustaining high emotional arousal and memory are linked, especially through the adrenal stress hormones, itself mediated by the amygdala activity (McGaugh, 2013).
In the study the dataset came from, Riegel et al. (2020) used words to elicit emotions (disgust, fear or neutral). Each pair displayed the same emotion and when showed together they had to imagine an interaction between them, forcing the creation of a meaningful association. To be sure that the displayed words showed the correct emotion, affective ratings were asked as controls. Each participant had to rate each pair on a Likert scale (1 to 7) of disgust and fear.
In this project, we are going to use part of this dataset to analyse and compare three statistical methods. We will compare (a) a simple logistic ordinal regression, (b) a mixed-effect model considering the subjects' repetition, and (c) a mixed-effect model considering the subjects' repetition and the random effects.


Data description
----------------
Here is a description of the different variables:

* **Subjects**: the subject's number in the experiment.

* **Wordpairs**: the word pairs presented.

* **Emotion**: the emotion elicited by the word pairs (disgust, fear or neutral).

* **Gender**: the gender of the subjects (man or woman).

* **Disgust**: a Likert scale (1 to 7) of the subjects’ emotional subjective feeling of disgust.


```{r, echo=FALSE}
#Creating a working dataset
#Giving nice names to variables

subjects<-Raw_data$ï..code
wordpairs<-Raw_data$wordpair
emotion<-Raw_data$emotion
gender<-Raw_data$sex
disgust<-Raw_data$disgust
fear<-Raw_data$fear

#Creating a dataset
Data<-data.frame(subjects,wordpairs,emotion,gender,disgust)

#Adapting the variables
Data$emotion<-as.factor(Data$emotion)
levels(Data$emotion) <- c("Disgust", "Fear","Neutral")
Data$gender<-as.factor(Data$gender)
levels(Data$gender) <- c("Women", "Men")
Data$disgust<-as.factor(Data$disgust)

```

Data exploration (Dataset)
----------------

```{r basic explor, comment = NA}
str(Data)
st(Data) #use "vtable"
```
As one can see in the table, the distribution of emotion through the word pairs and the gender is equivalent. The distribution of the disgust response has a floor effect, more than 80% of it being 1 and 2.


Data exploration (Graphs) 
-------------------------

```{r Likert representation, echo=FALSE}

## Prepare data <- **need to specify emotion$disgust + the other ratings**
Newdata<-data.frame(Data$emotion,Data$disgust)

df = prop.table(table(Newdata$Data.emotion, Newdata$Data.disgust), margin = 1)*100

#this is just ot make it fit into Likert
df = as.data.frame.matrix(df)
df$Item <- row.names(df)
rownames(df) = NULL
df = df[,c(8,1,2,3,4,5,6,7)]
colnames(df) = c("Item", "1 (low)","2","3","4","5","6","7 (high)")
```


```{r}
# Pretty Plot
plot(likert(summary = df))+ 
  guides(fill = guide_legend(name="Subjective evaluation"))+ 
  labs(title = "Subjective evaluation of disgust across all pairs of words")+
  labs(y = "Percentage of Likert answers") +
  labs(x="Word pairs' emotions") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.title = element_text(size = rel(1.5))) +
  theme( legend.box.background = element_rect(), legend.box.margin = margin(6, 6, 6, 6), legend.position = "right")
```

This figure shows the distribution of the subjective evaluation of disgust across all pair of words, depending of the emotion elicited by the words. As one can see, there still is a floor effect, but it seems less strong for the disgust emotion, as one can expect.


```{r alternate graph, include=FALSE}
##Disgust depending on emotion presented
ggplot(Data, aes(emotion,disgust,colour=gender))+geom_jitter()+
  labs(title = "Distribution of disgust evaluation", x = "Pairs of words' emotions", y = "Likert scale of subjective disgust feeling") + theme(plot.title = element_text(hjust = 0.75)) + labs(colour = "Gender")+ theme(plot.title = element_text(size = rel(2)))
```

Modelisation
============

Even if we have seen in our exploration that there seems to be a floor effect, we must do a diagnostic of a linear model to certify the inadaptation of this model to analyse our data.

```{r Basic analysis, include= FALSE}
disgust2<-as.numeric(disgust)
m0 = lm(disgust2~emotion*gender, data=Data)
plot(m0)

```

```{r normal analysis of resids}
sjPlot::plot_model(m0,type = "diag")[[3]]
```

As expected, the residuals distribution does not follow a normal curve. Therefore, we cannot do a classic linear regression to analyse this data. Alternative techniques have to be investigated. (Find more diagnostic graphs in the appendix.)

Three possible methods are being tested: (a) a simple logistic ordinal regression, (b) a mixed-effect model considering the subjects' repetition, and (c) a mixed-effect model considering the subjects' repetition and the random effect of the emotions. 

After modeling these methods, we can use the Akaike information criterion to estimate the prediction error of each model, and thereby their relative quality. 
```{r, include=FALSE}
#best method
m1.2 = clm(disgust ~ emotion * gender, data = Data)
m2.2 = clmm(disgust ~ emotion * gender + (1|subjects), data = Data)
m3.2 = clmm(disgust ~ emotion * gender + (emotion | subjects), data = Data)
a1<- extractAIC(m1.2)
a2<- extractAIC(m2.2)
a3<- extractAIC(m3.2)
```

```{r, echo=FALSE}
c1<-cbind(a1,a2,a3)
r1<-c1[1,]
r2<-c1[2,]
c2<-as.data.frame(cbind(r1,r2))
rownames(c2) <- c("Model A", "Model B","Model C")
colnames(c2) <- c("Equivalent degrees of freedom", "Akaike Information Criterion")
c2
```

As shown in this table, we can conclude that the model that seems to fit the best our data is the Model C, the mixed-effect model considering the subjects' repetition and the random effect of the emotions, as it has the smaller AIC, and more equivalent degrees of freedom.
This being done, we must now look for the best model using this method. According to our initial hypothesis, we want to know if there is an impact of the word pairs' emotion and the gender on the subjective evaluation of disgust. 

```{r best model, include=FALSE}
#best model
#m3.2 = clmm(disgust ~ emotion * gender + (emotion | subjects), data = Data) #done above
m3.1 = clmm(disgust ~ emotion + gender + (emotion | subjects), data = Data)
m3.0 = clmm(disgust ~ emotion + (emotion | subjects), data = Data)

a4<-extractAIC(m3.2)
a5<-extractAIC(m3.1)
a6<-extractAIC(m3.0)
```

```{r,echo=FALSE}
c3<-cbind(a4,a5,a6)
r3<-c3[1,]
r4<-c3[2,]
c4<-as.data.frame(cbind(r3,r4))
rownames(c4) <- c("Interaction Model", "Emotion and Gender Model","Model without Gender")
colnames(c4) <- c("Equivalent degrees of freedom", "Akaike Information Criterion")
c4
```
As shown in the table above, and following the AIC estimate, the model that fits the best to our data is the Model without Gender. 


Model analysis
==============
Now that we have choosen the Model without Gender as being the model that better fits our data, we have to analyse it.  

```{r best model analysis}
summary(m3.0)
(ctable <- coef(summary(m3.0)))

```

That means that the word pairs' emotion explains the disgust evaluation (p< 0.001). 

The estimated model can be written as:
$$Logit(P(Y\le 2))= (-0.5346) - (-2.0596)*emotionFear - (-3.9909)*emotionNeutral$$
$$Logit(P(Y\le 3))= 0.3387 - (-2.0596)*emotionFear - (-3.9909)*emotionNeutral$$
$$Logit(P(Y\le 4))= 1.1585 - (-2.0596)*emotionFear - (-3.9909)*emotionNeutral$$
$$Logit(P(Y\le 5))= 1.8831 - (-2.0596)*emotionFear - (-3.9909)*emotionNeutral$$
$$Logit(P(Y\le 6))= 2.6467 - (-2.0596)*emotionFear - (-3.9909)*emotionNeutral$$
$$Logit(P(Y\le 7))= 3.4641 - (-2.0596)*emotionFear - (-3.9909)*emotionNeutral$$

Being the model that best fits the data doesn't mean that it's a good model, that is why we must check for the effect size. 

```{r, echo=FALSE}
sjPlot::tab_model(m3.0)

```

In this table, we can see that the Conditional R^2 is 0.528, which means that the model explains 52,8% of the variance explained by both the fixed and random factors.  

Diagnostic of the model
=======================
```{r}

#read this : David W.. Hosmer, Lemeshow, S., & Rodney X.. Sturdivant. (2000). Applied logistic regression. New York: Wiley. --> §5

autoplot.clm(m1.2)
plot(m0, which = 2)
```

Conclusion
==========
The present data analysis was aimed to analyse and compare three statistical methods to analyse a dataset that 
