---
title: "Week11"
author: "Daniel Granja"
date: "08/12/2021"
output: 
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Aim
**Report how bone marrow transplant survival times relates to graft versus host disease (GHVD)**

#Preprocessing
##Load library and Data
```{r}
library(GGally)       # for ggpairs
library(ggfortify)    # for autoplot
library(ggplot2)      # for ggplot
library(ISwR)         # for the dataset

d<-graft.vs.host

#Select variables of interest
d[c(1,2,6)]<-NULL

```

##Data explanation
> This data aims to find the variables that explain bone marrow transplant survival times in relation with graft versus host disease (GHVD). 

```{r}
str(d)
summary(d)
```

Transform into factor the variables : type, preg, gvhd, dead

```{r}
d$type <-as.factor(d$type) #type of leukaemia coded 1: AML, 2: ALL, 3: CML for acute myeloid, acute lymphatic, and chronic myeloid leukaemia.
d$preg<-as.factor(d$preg) # indicating whether donor has been pregnant. 0: no, 1: yes.
levels(d$preg)<-c("no","yes")
d$gvhd<-as.factor(d$gvhd)# graft-versus-host disease, 0: no, 1: yes
levels(d$gvhd)<-c("no","yes")
d$dead <- as.factor(d$dead) # a numeric vector code, 0: no (censored), 1: yes
levels(d$dead)<-c("no","yes")
str(d)
summary(d)
```

#Plot the ggpairs
```{r}
ggpairs(d, aes(color=gvhd, alpha = 0.3))
```

#Modelization
```{r}
m1<-lm(time~ gvhd,data=d)
summary(m1)
```
> The p-value is significant for this model

#Adding other variables

```{r}
m2<-lm( time~ gvhd+preg, data=d)
summary(m2)
anova(m1,m2)
m3<-lm( time~ gvhd+type, data=d)
summary(m3)
anova(m1,m3)
m4<-lm( time~ gvhd+donage, data=d)
summary(m4)
anova(m1,m4)
```
> Only the addition of the type IV brings something more to the model as per the anova tests' significance. 

#Adding interaction
```{r}
m31<-lm( time~ gvhd*type, data=d)
summary(m31)
anova(m3,m31)
```
> The interaction doesn't bring much (p =< 0.75).
> The best model is m3: " Y = 747.6 + (-636.7)*gvhdyes + 145.8*type2 + 561.2*type3 ".

#Controling the postulates
```{r}
autoplot(m3)
```
> The postulates seem ok, even if we can discuss some bortherline observations.

#Representation
```{r}
ggplot(data=d, aes(x=gvhd,y=time, colour=type))+geom_violin()
```

>In conclusion, we can predict that the graft-versus-host desease reduce survival time when it's present, and the acute lymphatic leukaemia (type 2) and the chronic myeloid leukaemia (type 3) are associated with a better survival time than acute myeloid leukaemia.

# Version of R used 
```{r, echo = FALSE}
sessionInfo()
```